\documentclass[11pt,a4paper]{article}
\input{mypackages}
\input{mystyle}
\input{sections/glossary}
\graphicspath{{figs/}}
\newcounter{myCounter}

\begin{document}

\title{\ipcFull (\ipc)\thanks{\url{https://ipc.space}}}
\date{}
\author{ConsensusLab\thanks{\url{https://consensuslab.world}}}
% \authornote{Both authors contributed equally to this research.}
% \orcid{1234-5678-9012}
% \author{G.K.M. Tobin}
% \authornotemark[1]
%\email{sgoren@campus.technion.ac.il}
% \orcid{https://orcid.org/0000-0003-2158-161X}
% \affiliation{%
%     \institution{Technion}
%     \country{Israel}
%     }

\maketitle
\begin{abstract}
Totally ordering transactions constitutes a
significant scalability bottleneck standing in the way of massive adoption of
blockchain systems, as all participants need to replicate and execute sequentially every single transaction. Layer 2 (L2) protocols aim at resolving this scalability
limitation by off-loading state and processing to a loosely coupled sub-system.

We present \ipcFull (\ipc), a new blockchain
architecture design that enhances the scaling capabilities of L2+
protocols. Users of \ipc can dynamically spawn new blockchain subsystems
(subnets) as children of any existing subnet.  \ipc is based on the design principles of
on-demand horizontal scaling. Child subnets leverage the
security of their parent subnets by periodically checkpointing their state in the parent's state.
\ipc provides native communication across
subnets within the \ipc framework.

We believe that \ipc addresses several issues with existing L2 approaches,
including limited throughput capacity, isolation from each
other, centralized components, or monolithic architectures.
% We believe, however, that \ipc's flexible architecture could be implemented in
% conjunction with many of these orthogonal solutions.
In the following, we introduce the overall system architecture, native functionality,
and design decisions of our reference implementation for child
subnets based on the BFT Trantor consensus protocol, with Filecoin as
rootnet.
\end{abstract}

% ----------------------------------------------------------------
% ----------------------------------------------------------------

\input{sections/introduction}
\input{sections/example-use-case}
\input{sections/preliminaries}
\input{sections/functionality}
\input{sections/components}
\input{sections/related-work}
\input{sections/ref-impl}
\input{sections/conclusion}
 
% ----------------------------------------------------------------
% ----------------------------------------------------------------

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{plain}
\bibliography{bibliography}

%%
%% If your work has an appendix, this is the place to put it.
\appendix
% ----------------------------------------------------------------
% ----------------------------------------------------------------
\newpage
\printglossaries
% \input{sections/appendix.tex}
% \input{sections/algorithm.tex}
% \input{sections/QvW.tex}
% \input{sections/Yao.tex}
% ----------------------------------------------------------------
% ----------------------------------------------------------------

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}
\endinput
%%
%% End of file 
